Distributions of Random Variables

### 2018-09-29 DATA606/Distributions of Random Variables .Rmd
More work

---
title: "Distributions of Random Variables"
author: "Mike Silva"
date: "September 20, 2018"
output: html_document
---

@@

(d) What percent of the triathletes did Leo finish faster than in his group?

@@

(e) What percent of the triathletes did Mary finish faster than in her group?

@@

(a) The mean height is 61.52 inches with a standard deviation of 4.58 inches. Use this information to determine if the heights approximately follow the 68-95-99.7% Rule.

@@

**Yes, they roughly follow the 68-95-99.7 Rule.**

(b) Do these data appear to follow a normal distribution? Explain your reasoning using the graphs provided below.

@@

**Yes.  The dots on the Q-Q plot roughly fall on the line.  This indicates the data are normally distributed.**

@@

A machine that produces a special type of transistor (a component of computers) has a 2% defective rate. The production is considered a random process where each transistor is independent of the others.
(a) What is the probability that the 10th transistor produced is the first with a defect?
(b) What is the probability that the machine produces no defective transistors in a batch of 100?
(c) On average, how many transistors would you expect to be produced before the first with a defect? What is the standard deviation?
(d) Another machine that also produces transistors has a 5% defective rate where each transistor is produced independent of the others. On average how many transistors would you expect to be produced with this machine before the first with a defect? What is the standard deviation?
(e) Based on your answers to parts (c) and (d), how does increasing the probability of an event affect the mean and standard deviation of the wait time until success?

@@

While it is often assumed that the probabilities of having a boy or a girl are the same, the actual probability of having a boy is slightly higher at 0.51. Suppose a couple plans to have 3 kids.
(a) Use the binomial model to calculate the probability that two of them will be boys.
(b) Write out all possible orderings of 3 children, 2 of whom are boys. Use these scenarios to calculate the same probability from part (a) but using the addition rule for disjoint outcomes. Confirm that your answers from parts (a) and (b) match.
(c) If we wanted to calculate the probability that a couple who plans to have 8 kids will have 3 boys, briefly describe why the approach from part (b) would be more tedious than the approach from part (a).

## 3.42 2 Serving in Volleyball

@@

(a) What is the probability that on the 10th try she will make her 3rd successful serve?
(b) Suppose she has made two successful serves in nine attempts. What is the probability that her 10th serve will be successful?
(c) Even though parts (a) and (b) discuss the same scenario, the probabilities you calculated should be different. Can you explain the reason for this discrepancy?

>>>>

---
title: "Distributions of Random Variables"
author: "Mike Silva"
date: "September 20, 2018"
output:
  html_document:
    theme: yeti
    highlight: tango
---

@@

(d) What percent of the triathletes did Leo finish faster than in his group? **About 14%**

@@

(e) What percent of the triathletes did Mary finish faster than in her group? **Roughly 38%**

@@

(a) The mean height is 61.52 inches with a standard deviation of 4.58 inches. Use this information to determine if the heights approximately follow the 68-95-99.7% Rule. **Yes, they roughly follow the 68-95-99.7 Rule.**

@@

(b) Do these data appear to follow a normal distribution? Explain your reasoning using the graphs provided below. **Yes.  The dots on the Q-Q plot roughly fall on the line.  This indicates the data are normally distributed.**

@@

@@

A machine that produces a special type of transistor (a component of computers) has a 2% defective rate. The production is considered a random process where each transistor is independent of the others.

(a) What is the probability that the 10th transistor produced is the first with a defect? **The probability is roughly 0.017.**

```{r}
probability_of_defect <- 0.02

probability_of_10th_first_defect = 1

for(i in 1:9){
  probability_of_10th_first_defect <- probability_of_10th_first_defect * (1 - probability_of_defect)
}

probability_of_10th_first_defect <- probability_of_10th_first_defect * probability_of_defect

probability_of_10th_first_defect
```

This matches the geometric distribution formula's 
$(1-p)^{n-1}\times p$

```{r}
((1 - 0.02)^(10-1))*0.02
```

(b) What is the probability that the machine produces no defective transistors in a batch of 100? **About 0.13.**

```{r}
probability_of_no_defect = 1

for(i in 1:100){
   probability_of_no_defect <- probability_of_no_defect * (1 - probability_of_defect)
}

probability_of_no_defect
```

(c) On average, how many transistors would you expect to be produced before the first with a defect? **A 2% defect rate means out of 100 produced 2 would be defective.  Therefore we would expect 1 defective transitor out of 50.**  What is the standard deviation? **0.14**

```{r}
sigma <- sqrt(probability_of_defect*(1 - probability_of_defect))
sigma
```

(d) Another machine that also produces transistors has a 5% defective rate where each transistor is produced independent of the others. On average how many transistors would you expect to be produced with this machine before the first with a defect? **Using the same logice I would expect 20**  What is the standard deviation? **Apx 0.22**

```{r}
sigma <- sqrt(0.05*(1 - 0.05))
sigma
```

(e) Based on your answers to parts (c) and (d), how does increasing the probability of an event affect the mean and standard deviation of the wait time until success?  **It decreases the mean and increases the standard deviation.**

@@

While it is often assumed that the probabilities of having a boy or a girl are the same, the actual probability of having a boy is slightly higher at 0.51. Suppose a couple plans to have 3 kids.

(a) Use the binomial model to calculate the probability that two of them will be boys. **0.38**

```{r}
k <- 2
n <- 3
p <- 0.51
((factorial(n)/(factorial(k)*factorial(n-k))) * p^k)*(1-p)^(n-k)
```

(b) Write out all possible orderings of 3 children, 2 of whom are boys. Use these scenarios to calculate the same probability from part (a) but using the addition rule for disjoint outcomes. Confirm that your answers from parts (a) and (b) match. **The combinations that will produce this kind of outcome are: girl-boy-boy, boy-girl-boy, boy-boy-girl**

```{r}
p_gbb <- (1 - p) * p * p
p_bgb <- p * (1 - p) * p
p_bbg <- p * p * (1 - p)

p_gbb + p_gbb + p_bbg
```

(c) If we wanted to calculate the probability that a couple who plans to have 8 kids will have 3 boys, briefly describe why the approach from part (b) would be more tedious than the approach from part (a).  **There would be more combinations that would need to have probabilities computed which would be more tedious.**

## 3.42 Serving in Volleyball

@@

(a) What is the probability that on the 10th try she will make her 3rd successful serve?  **The chances that her 10th being her 3rd success is roughly 0.04.**

```{r}
# Compute the chances of having 2 successful serve out of 9
k <- 2
n <- 9
p <- 0.15
p_2_out_of_9 <- ((factorial(n)/(factorial(k)*factorial(n-k))) * p^k)*(1-p)^(n-k)
# Now factor in the chance of her 10th being a success
p_2_out_of_9 * p
```

(b) Suppose she has made two successful serves in nine attempts. What is the probability that her 10th serve will be successful?  **15% because each serve is independent of the others.**

(c) Even though parts (a) and (b) discuss the same scenario, the probabilities you calculated should be different. Can you explain the reason for this discrepancy?  **The probability in (b) is for a single event, while the probability in (a) is for a series of events.  That's why they are different.**
